Matteo Riondato was born in Padua, Italy, in January 1986. He attended
Universit\`a degli Studi di Padova, where he obtained a Laurea (Sc.B.) in Information
Engineering with a honors thesis on ``Algorithmic Aspects of Cryptography'' with
Prof.~Andrea Pietracaprina. He then obtain a Laurea Specialistica (Sc.M)
\emph{cum laude} in Computer Enginering with a thesis on ``Top-K Frequent
Itemsets Mining through Sampling'', having Prof.~Andrea Pietracaprina, Prof.~Eli
Upfal, and Dr.~Fabio Vandin as advisor. He spent the last year of the master as
a visiting student at the Department of Computer Science at Brown University. He
joined the Ph.D.~program in computer science at Brown in Fall 2009, with
Prof.~Eli Upfal as advisor. He was a member of both the theory group and the
data management group. He spent summer 2010 as a visiting student at Chalmers
University, Gothemburg, Sweden; summer 2011 as a research fellow at the
Department of Information Engineering at the Universit\`a degli Studi di Padova,
Padua, Italy; summer 2012 as a visiting student at Sapienza University, Rome,
Italy; and summer 2013 as a research intern at Yahoo Research Barcelona. While
at Brown, he was the teaching assistant for CSCI 1550 (Probabilistic Methods in
Computer Science / Probability and Computing) for four times. He was also
President of the Graduate Student Council (Apr 2011 -- Dec 2012), the official
organization of graduate students at Brown. He was a student representative on
the Graduate Council (2011 -- 2013) and on the Presidential Strategic Planning
Committee for Doctoral Education.

