\begin{thebibliography}{42}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Abraham et~al.(2011)Abraham, Delling, Fiat, Goldberg, and
  Werneck]{AbrahamDFGW11}
I.~Abraham, D.~Delling, A.~Fiat, A.~V. Goldberg, and R.~F. Werneck.
\newblock {VC}-dimension and shortest path algorithms.
\newblock In \emph{Automata, Languages and Programming}, volume 6755 of
  \emph{Lecture Notes in Computer Science}, pages 690--699, 2011.
%\newblock \doi{10.1007/978-3-642-22006-7_58}.

\bibitem[Alon and Spencer(2008)]{AlonS08}
N.~Alon and J.~H. Spencer.
\newblock \emph{The Probabilistic Method}.
\newblock Interscience Series in Discrete Mathematics and Optimization. John
  Wiley {\&} Sons, Hoboken, NJ, USA, third edition, 2008.

\bibitem[Bartlett et~al.(2002)Bartlett, Boucheron, and Lugosi]{BartlettBL02}
P.~L. Bartlett, S.~Boucheron, and G.~Lugosi.
\newblock Model selection and error estimation.
\newblock \emph{Machine Learning}, 48:\penalty0 85--113, 2002.
%\newblock \doi{10.1023/A:1013999503812}.

\bibitem[Benjamini and Hochberg(1995)]{BenjaminiH95}
Y.~Benjamini and Y.~Hochberg.
\newblock Controlling the false discovery rate: A practical and powerful
  approach to multiple testing.
\newblock \emph{Journal of the Royal Statistical Society. Series B
  (Methodological)}, 57\penalty0 (1):\penalty0 289--300, 1995.
%\newblock ISSN 00359246.
%\newblock URL \url{http://www.jstor.org/stable/2346101}.

\bibitem[Blumer et~al.(1989)Blumer, Ehrenfeucht, Haussler, and
  Warmuth]{BlumerEHW89}
A.~Blumer, A.~Ehrenfeucht, D.~Haussler, and M.~K. Warmuth.
\newblock Learnability and the {V}apnik-{C}hervonenkis dimension.
\newblock \emph{Journal of the ACM}, 36:\penalty0 929--965, October 1989.
%\newblock ISSN 0004-5411.
%\newblock \doi{http://doi.acm.org/10.1145/76359.76371}.
%\newblock URL \url{http://doi.acm.org/10.1145/76359.76371}.

\bibitem[Bolton et~al.(2002)Bolton, Hand, and Adams]{BoltonHA02}
R.~J. Bolton, D.~J. Hand, and N.~M. Adams.
\newblock Determining hit rate in pattern search.
\newblock In \emph{Pattern Detection and Discovery}, volume 2447 of
  \emph{Lecture Notes in Computer Science}, pages 36--48, 2002.
%\newblock \doi{10.1007/3-540-45728-3_4}.

\bibitem[Boucheron et~al.(2005)Boucheron, Bousquet, and Lugosi]{BoucheronBL05}
S.~Boucheron, O.~Bousquet, and G.~Lugosi.
\newblock Theory of classification : A survey of some recent advances.
\newblock \emph{{ESAIM}: Probability and Statistics}, 9:\penalty0 323--375,
  2005.
%\newblock URL
%  \url{http://cat.inist.fr/?aModele=afficheN\&\#38;cpsidt=17367966}.

\bibitem[Brijs et~al.(1999)Brijs, Swinnen, Vanhoof, and Wets]{BrijsSVW99}
T.~Brijs, G.~Swinnen, K.~Vanhoof, and G.~Wets.
\newblock Using association rules for product assortment decisions: A case
  study.
\newblock In U.~M. Fayyad, S.~Chaudhuri, and D.~Madigan, editors, \emph{KDD},
  pages 254--260. ACM, 1999.
%\newblock ISBN 1-58113-143-7.

\bibitem[Calders et~al.(2006)Calders, Rigotti, and Boulicaut]{CaldersRB06}
T.~Calders, C.~Rigotti, and J.-F. Boulicaut.
\newblock A survey on condensed representations for frequent sets.
\newblock In \emph{Constraint-Based Mining and Inductive Databases}, volume
  3848 of \emph{Lecture Notes in Computer Science}, pages 64--80, 2006.
%\newblock \doi{10.1007/11615576_4}.

\bibitem[Chakaravarthy et~al.(2009)Chakaravarthy, Pandit, and
  Sabharwal]{ChakaravarthyPS09}
V.~T. Chakaravarthy, V.~Pandit, and Y.~Sabharwal.
\newblock Analysis of sampling techniques for association rule mining.
\newblock In \emph{Proceedings of the 12th International Conference on Database
  Theory}, ICDT '09, pages 276--283, New York, NY, USA, 2009. ACM.
%\newblock ISBN 978-1-60558-423-2.
%\newblock \doi{http://doi.acm.org/10.1145/1514894.1514927}.
%\newblock URL \url{http://doi.acm.org/10.1145/1514894.1514927}.

\bibitem[Chazelle(2000)]{Chazelle00}
B.~Chazelle.
\newblock \emph{The discrepancy method: randomness and complexity}.
\newblock Cambridge University Press, New York, NY, USA, 2000.
%\newblock ISBN 0-521-00357-1.

\bibitem[Devroye et~al.(1996)Devroye, Gy{\"o}rfi, and Lugosi]{DevroyeGL96}
L.~Devroye, L.~Gy{\"o}rfi, and G.~Lugosi.
\newblock \emph{A Probabilistic Theory of Pattern Recognition}.
\newblock Applications of Mathematics. Springer, 1996.
%\newblock ISBN 9780387946184.
%\newblock URL \url{http://books.google.com/books?id=uDgXoRkyWqQC}.

\bibitem[Dudoit et~al.(2003)Dudoit, Shaffer, and Boldrick]{DudoitSB03}
S.~Dudoit, J.~Shaffer, and J.~Boldrick.
\newblock Multiple hypothesis testing in microarray experiments.
\newblock \emph{Statistical Science}, 18\penalty0 (1):\penalty0 71--103, 2003.
%\newblock ISSN 08834237.
%\newblock \doi{10.2307/3182872}.

\bibitem[DuMouchel and Pregibon(2001)]{DuMouchelP01}
W.~DuMouchel and D.~Pregibon.
\newblock Empirical {B}ayes screening for multi-item associations.
\newblock In \emph{Proceedings of the seventh ACM SIGKDD international
  conference on Knowledge discovery and data mining}, KDD '01, pages 67--76,
  New York, NY, USA, 2001. ACM.
%\newblock ISBN 1-58113-391-X.
%\newblock \doi{10.1145/502512.502526}.
%\newblock URL \url{http://doi.acm.org/10.1145/502512.502526}.

\bibitem[Geng and Hamilton(2006)]{GengH06}
L.~Geng and H.~J. Hamilton.
\newblock Interestingness measures for data mining: A survey.
\newblock \emph{ACM Computing Surveys}, 38\penalty0 (3), Sept. 2006.
%\newblock ISSN 0360-0300.
%\newblock \doi{10.1145/1132960.1132963}.
%\newblock URL \url{http://doi.acm.org/10.1145/1132960.1132963}.

\bibitem[Geurts et~al.(2003)Geurts, Wets, Brijs, and Vanhoof]{GeurtsWBV03}
K.~Geurts, G.~Wets, T.~Brijs, and K.~Vanhoof.
\newblock Profiling of high-frequency accident locations by use of association
  rules.
\newblock \emph{Transportation Research Record: Journal of the Transportation
  Research Board}, 1840:\penalty0 123--130, 2003.

\bibitem[Gionis et~al.(2007)Gionis, Mannila, Mielik\"{a}inen, and
  Tsaparas]{GionisMMT07}
A.~Gionis, H.~Mannila, T.~Mielik\"{a}inen, and P.~Tsaparas.
\newblock Assessing data mining results via swap randomization.
\newblock \emph{ACM Transactions on Knowledge Discovery from Data}, 1\penalty0
  (3), Dec. 2007.
%\newblock ISSN 1556-4681.
%\newblock \doi{10.1145/1297332.1297338}.
%\newblock URL \url{http://doi.acm.org/10.1145/1297332.1297338}.

\bibitem{Goethals:2004:AFI:1007730.1007744}
B.~Goethals and M.~J. Zaki.
\newblock Advances in frequent itemset mining implementations: report on
  FIMI'03.
\newblock {\em SIGKDD Explor. Newsl.}, 6(1):109--117, June 2004.

\bibitem[Goldshmidt et~al.(1994)Goldshmidt, Nehme, and Yu]{GoldschmidtNY94}
O.~Goldshmidt, D.~Nehme, and G.~Yu.
\newblock Note: On the set-union knapsack problem.
\newblock \emph{Naval Research Logistics}, 41\penalty0 (6):\penalty0 833--842,
  1994.

\bibitem[Grahne and Zhu(2003)]{GrahneZ03}
G.~Grahne and J.~Zhu.
\newblock Efficiently using prefix-trees in mining frequent itemsets.
\newblock In B.~Goethals and M.~J. Zaki, editors, \emph{FIMI}, volume~90 of
  \emph{CEUR Workshop Proceedings}. CEUR-WS.org, 2003.

\bibitem[H{\"a}m{\"a}l{\"a}inen(2010)]{Hamalainen10}
W.~H{\"a}m{\"a}l{\"a}inen.
\newblock {StatApriori}: an efficient algorithm for searching statistically
  significant association rules.
\newblock \emph{Knowledge and Information Systems}, 23\penalty0 (3):\penalty0
  373--399, 2010.
%\newblock \doi{10.1007/s10115-009-0229-8}.

\bibitem[Han et~al.(2007)Han, Cheng, Xin, and Yan]{HanCXY07}
J.~Han, H.~Cheng, D.~Xin, and X.~Yan.
\newblock Frequent pattern mining: current status and future directions.
\newblock \emph{Data Mining and Knowledge Discovery}, 15:\penalty0 55--86,
  2007.
%\newblock ISSN 1384-5810.
%\newblock \doi{10.1007/s10618-006-0059-1}.
%\newblock URL \url{http://dx.doi.org/10.1007/s10618-006-0059-1}.

\bibitem[Hanhij{\"a}rvi(2011)]{Hanhijarvi11}
S.~Hanhij{\"a}rvi.
\newblock Multiple hypothesis testing in pattern discovery.
\newblock In \emph{Discovery Science}, volume 6926 of \emph{Lecture Notes in
  Computer Science}, pages 122--134, 2011.
%\newblock \doi{10.1007/978-3-642-24477-3_12}.

\bibitem[Har-Peled and Sharir(2011)]{HarPS11}
S.~Har-Peled and M.~Sharir.
\newblock Relative $(p,\varepsilon)$-approximations in geometry.
\newblock \emph{Discrete \& Computational Geometry}, 45\penalty0 (3):\penalty0
  462--496, 2011.
%\newblock ISSN 0179-5376.
%\newblock URL \url{http://dx.doi.org/10.1007/s00454-010-9248-1}.

\bibitem[Holm(1979)]{Holm79}
S.~Holm.
\newblock A simple sequentially rejective multiple test procedure.
\newblock \emph{Scandinavian Journal of Statistics}, 6\penalty0 (2):\penalty0
  65--70, 1979.
%\newblock ISSN 03036898.
%\newblock URL \url{http://www.jstor.org/stable/4615733}.

\bibitem[Jacquemont et~al.(2009)Jacquemont, Jacquenet, and
  Sebban]{JacquemontFS09}
S.~Jacquemont, F.~Jacquenet, and M.~Sebban.
\newblock A lower bound on the sample size needed to perform a significant
  frequent pattern mining task.
\newblock \emph{Pattern Recognition Letters}, 30\penalty0 (11):\penalty0
  960--967, 2009.
%\newblock ISSN 0167-8655.
%\newblock \doi{10.1016/j.patrec.2009.05.002}.
%\newblock URL
%  \url{http://www.sciencedirect.com/science/article/pii/S0167865509000968}.

\bibitem[Jaroszewicz et~al.(2009)Jaroszewicz, Scheffer, and
  Simovici]{JaroszewiczSS09}
S.~Jaroszewicz, T.~Scheffer, and D.~A. Simovici.
\newblock Scalable pattern mining with bayesian networks as background
  knowledge.
\newblock \emph{Data Mining and Knowledge Discovery}, 18\penalty0 (1):\penalty0
  56--100, 2009.
%\newblock \doi{10.1007/s10618-008-0102-5}.

\bibitem[Kirsch et~al.(2012)Kirsch, Mitzenmacher, Pietracaprina, Pucci, Upfal,
  and Vandin]{KirschMAPUV12}
A.~Kirsch, M.~Mitzenmacher, A.~Pietracaprina, G.~Pucci, E.~Upfal, and
  F.~Vandin.
\newblock An efficient rigorous approach for identifying statistically
  significant frequent itemsets.
\newblock \emph{Journal of the ACM}, 59\penalty0 (3):\penalty0 12:1--12:22,
  June 2012.
%\newblock ISSN 0004-5411.
%\newblock \doi{10.1145/2220357.2220359}.
%\newblock URL \url{http://doi.acm.org/10.1145/2220357.2220359}.

\bibitem[Kleinberg et~al.(2004)Kleinberg, Sandler, and Slivkins]{KleinbergSS04}
J.~Kleinberg, M.~Sandler, and A.~Slivkins.
\newblock Network failure detection and graph connectivity.
\newblock In \emph{Proceedings of the fifteenth annual ACM-SIAM symposium on
  Discrete algorithms}, SODA '04, pages 76--85, Philadelphia, PA, USA, 2004.
  Society for Industrial and Applied Mathematics.
%\newblock ISBN 0-89871-558-X.
%\newblock URL \url{http://dl.acm.org/citation.cfm?id=982792.982803}.

\bibitem[Lallich et~al.(2006)Lallich, Teytaud, and Prudhomme]{LallichTP06}
S.~Lallich, O.~Teytaud, and E.~Prudhomme.
\newblock Statistical inference and data mining: false discoveries control.
\newblock In \emph{Compstat 2006 - Proceedings in Computational Statistics},
  pages 325--336, 2006.
%\newblock \doi{10.1007/978-3-7908-1709-6_25}.

\bibitem[Li et~al.(2001)Li, Long, and Srinivasan]{LiLS01}
Y.~Li, P.~M. Long, and A.~Srinivasan.
\newblock Improved bounds on the sample complexity of learning.
\newblock \emph{Journal of Computer and System Sciences}, 62\penalty0
  (3):\penalty0 516--527, 2001.
%\newblock ISSN 0022-0000.
%\newblock \doi{DOI: 10.1006/jcss.2000.1741}.
%\newblock URL
%  \url{http://www.sciencedirect.com/science/article/B6WJ0-45B65MX-H/2/3b0f2dd50a30dcee8b995d701959f15f}.

\bibitem[Liu et~al.(2011)Liu, Zhang, and Wong]{LiuZW11}
G.~Liu, H.~Zhang, and L.~Wong.
\newblock Controlling false positives in association rule mining.
\newblock \emph{Proceedings of the VLDB Endowement}, 5\penalty0 (2):\penalty0
  145--156, Oct. 2011.
%\newblock ISSN 2150-8097.
%\newblock URL \url{http://dl.acm.org/citation.cfm?id=2078324.2078330}.

\bibitem[L\"{o}ffler and Phillips(2009)]{LofflerP09}
M.~L\"{o}ffler and J.~M. Phillips.
\newblock Shape fitting on point sets with probability distributions.
\newblock In A.~Fiat and P.~Sanders, editors, \emph{Algorithms - ESA 2009},
  volume 5757 of \emph{Lecture Notes in Computer Science}, pages 313--324.
  Springer, Berlin / Heidelberg, 2009.
%\newblock URL \url{http://dx.doi.org/10.1007/978-3-642-04128-0_29}.

\bibitem[Matou\v{s}ek(2002)]{Matousek02}
J.~Matou\v{s}ek.
\newblock \emph{Lectures on Discrete Geometry}.
\newblock Springer-Verlag, Secaucus, NJ, USA, 2002.
%\newblock ISBN 0387953744.

\bibitem[Megiddo and Srikant(1998)]{MegiddoS98}
N.~Megiddo and R.~Srikant.
\newblock Discovering predictive association rules.
\newblock In R.~Agrawal, P.~E. Stolorz, and G.~Piatetsky-Shapiro, editors,
  \emph{Proceedings of the Fourth International Conference on Knowledge
  Discovery and Data Mining (KDD'98)}, pages 274--278, New York, NY, USA, 1998.
  AAAI Press.

\bibitem[Riondato and Upfal(2012)]{RiondatoU12}
M.~Riondato and E.~Upfal.
\newblock Efficient discovery of association rules and frequent itemsets
  through sampling with tight performance guarantees.
\newblock In P.~A. Flach, T.~De~Bie, and N.~Cristianini, editors, \emph{Machine
  Learning and Knowledge Discovery in Databases - European Conference, ECML
  PKDD 2012, Bristol, UK, September 24-28, 2012. Proceedings, Part I}, volume
  7523 of \emph{Lecture Notes in Computer Science}, pages 25--41. Springer,
  2012.
%\newblock ISBN 978-3-642-33459-7.

\bibitem[Riondato et~al.(2011)Riondato, Akdere, \c{C}etintemel, Zdonik, and
  Upfal]{RiondatoACZU11}
M.~Riondato, M.~Akdere, U.~\c{C}etintemel, S.~B. Zdonik, and E.~Upfal.
\newblock The {VC}-dimension of {SQL} queries and selectivity estimation
  through sampling.
\newblock In D.~Gunopulos, T.~Hofmann, D.~Malerba, and M.~Vazirgiannis,
  editors, \emph{Machine Learning and Knowledge Discovery in Databases -
  European Conference, ECML PKDD 2011, Athens, Greece, September 5-9, 2011.
  Proceedings, Part II}, volume 6912 of \emph{Lecture Notes in Computer
  Science}, pages 661--676. Springer, 2011.
%\newblock ISBN 978-3-642-23782-9.

\balance

\bibitem[Silverstein et~al.(1998)Silverstein, Brin, and
  Motwani]{SilversteinBM98}
C.~Silverstein, S.~Brin, and R.~Motwani.
\newblock Beyond market baskets: Generalizing association rules to dependence
  rules.
\newblock \emph{Data Mining and Knowledge Discovery}, 2\penalty0 (1):\penalty0
  39--68, Jan. 1998.
%\newblock ISSN 1384-5810.
%\newblock \doi{10.1023/A:1009713703947}.
%\newblock URL \url{http://dx.doi.org/10.1023/A:1009713703947}.

\bibitem[Teytaud and Lallich(2001)]{TeytaudL01}
O.~Teytaud and S.~Lallich.
\newblock Contribution of statistical learning to validation of association
  rules.
\newblock Extended version of "Bornes unformes en extraction de r\`{e}gles
  d'association", proceedings of CAp 2001, 2001.

\bibitem[Vapnik(1999)]{Vapnik99}
V.~N. Vapnik.
\newblock \emph{The Nature of Statistical Learning Theory}.
\newblock Statistics for engineering and information science. Springer-Verlag,
  New York, NY, USA, 1999.
%\newblock ISBN 9780387987804.

\bibitem[Vapnik and Chervonenkis(1971)]{VapnikC71}
V.~N. Vapnik and A.~J. Chervonenkis.
\newblock On the uniform convergence of relative frequencies of events to their
  probabilities.
\newblock \emph{Theory of Probability and its Applications}, 16\penalty0
  (2):\penalty0 264--280, 1971.
%\newblock \doi{10.1137/1116025}.

\bibitem[Webb(2007)]{Webb07}
G.~I. Webb.
\newblock Discovering significant patterns.
\newblock \emph{Machine Learning}, 68\penalty0 (1):\penalty0 1--33, 2007.
%\newblock \doi{10.1007/s10994-007-5006-x}.

\bibitem[Webb(2008)]{Webb08}
G.~I. Webb.
\newblock Layered critical values: a powerful direct-adjustment approach to
  discovering significant patterns.
\newblock \emph{Machine Learning}, 71:\penalty0 307--323, 2008.
%\newblock \doi{10.1007/s10994-008-5046-x}.

\end{thebibliography}

