\section{Conclusions}\label{sec:concl}

In this work we consider the problem of mining Real Frequent Itemsets (RFI's), 
that are itemsets whose probability of appearing in a transaction sampled from a
distribution is at least $\theta$, where $\theta$ is a user-specified parameter,
given only a transactional dataset that is a collection of independent
identically distributed samples from the distribution. Since every transactional 
dataset represents only a finite observation from the process for which
knowledge is inferred using the mining task, this is a fundamental
problem in data mining, but has received scant attention in the literature.

In this paper we present an algorithm to mine a collection of RFI's while providing guarantees on the quality of the returned collection. 
In particular, our method
guarantees that the probability of any spurious discovery (i.e., an itemset in the collection whose probability is below $\theta$)
 is within the user-specified parameter $\delta$. Our method is orthogonal to the numerous methods that have been  previously proposed in the literature to control false discoveries
 among the frequent itemsets, since such methods focus on the itemsets that are frequent in the dataset, assuming that they reflect the RFI's.
 
To identify a high quality collection of real frequent itemsets, we use results from statistical learning theory. We define a range set associated to the problem of mining RFI's,  and give an
upper bound to its (empirical) VC-dimension, showing an interesting connection
with a variant of the knapsack optimization problem.
To the best of our knowledge, ours is the first work to apply techniques from statistical learning theory to the field of RFI's, and in general the first application of the sample complexity
bound based on empirical VC-Dimension to the field of data mining. We
implemented our test and evaluated its performances on large datasets generated
from available transactional datasets. We found that it is very efficient on
multiple metrics. Our method can control the FWER even better than what is
guaranteed by the theoretical analysis. We evaluated its statistical power and
compared it to the power of other available statistical tests adapted to the
discovery of RFI's, and found it not only very high in absolute terms, but also
comparable or even better than the power of state-of-the-art tests.

There are a number of directions for further research, including how to improve the power of methods to identify RFI's, and studying lower bounds to the VC-dimension of the range set we define to mine RFI's.
Moreover, while this work focuses on itemsets mining, further research is needed for the extraction of more structured (e.g., sequences, graphs) real frequent patterns.
 

